Computing properties of independent sets by generic programming tensor networks \thanks{\funding{...}}

### 2021-09-12 paper/paper.tex
update paper

\begin{abstract}
We introduce a new approach to compute various properties of independent sets by encoding the problems and the graph into tensor networks with different tensor element algebra.
The type of graph problems that can be computed using this method includes: independence number, the number of independent sets, independence polynomial, maximal independence polynomial, and both counting and enumeration of all maximum independence sets, the large-size independent sets, and all maximal independent sets.
Although the computational complexity inevitably scales exponentially with the graph size, our method using tensor network contraction with generic programming is highly versatile and is able to compute properties that are otherwise not known to be possible for up to several hundred vertices.
At the same time, our algorithms are very simple to implement and one can directly utilize recent advances in tensor network contraction techniques.
To demonstrate the versatility of this tool, we apply it to a few examples including the calculations of the hard-square entropy constant, Euler characteristics of the independence complex, partition functions, and finite-temperature phase transitions on square-lattice graphs.
\end{abstract}

@@

The independent set problem is closely related to the clique problem and the vertex cover problem~\cite{Moore2011};
more concretely, an MIS of a graph $G$ corresponds to a maximum clique in the complement graph of $G$,
and the complement vertex set of an MIS corresponds to a minimum vertex cover of the graph $G$.
Together, these problems find a wide range of applications in scheduling, logistics, social network analysis, bioinformatics, wireless networks and telecommunication,
map labelling, computer vision, etc.~\cite{Butenko2003, Wu2015}.
Besides the standard problem of finding the independence number and an MIS, there are many other interesting problems pertaining to independent sets, such as the number of independent sets,
independence polynomial, maximal independence polynomial, and enumeration of MIS or maximal independent set configurations.
Some of these problems are of great interest in physics applications such as the hard-core lattice gas model~\cite{Dyre2016, Fernandes2007}
in statistical mechanics and the Rydberg hamiltonian with neutral atoms~\cite{Pichler2018} \red{cite experiment when it's ready};
they can, for example, be used to understand phase transitions and to identify harder graphs in an ensemble of graphs \red{cite experiment}.

@@

\begin{example}
$C_{ijk} = A_{jkm} B_{mil} V_{jm}$ is a tensor network representing $C_{ijk} = \sum_{ml}A_{jkm} B_{mil} V_{jm}$,
where $i,j,k$ in the output tensor are open edges (note that $j$ is not an open edge and thus summed over).
Its hypergraph representation is shown below, where we use different colors to annotate different hyperedges.

@@

This motivates us to think about what is the most general element type that is allowed in a tensor network contraction program.
We find that as long as the algebra of tensor elements forms a commutative semiring, the tensor network contraction result will be valid and be independent of the contraction order.
A commutative semiring is a semiring with its multiplication operation being commutative, while a semiring is a ring without additive inverse.
To define a commutative semiring with the addition operation $\oplus$ and the multiplication operation $\odot$ on a set $R$, the following relations must hold for any arbitrary three elements $a, b, c \in R$.
\begin{align*}
(a \oplus b) \oplus c = a \oplus (b \oplus c) & \hspace{5em}\text{$\triangleright$ commutative monoid $\oplus$ with identity $\mymathbb{0}$}\\
a \oplus \mymathbb{0} = \mymathbb{0} \oplus a = a &\\
a \oplus b = b \oplus a &\\
&\\
(a \odot b) \odot c = a \odot (b \odot c)  &   \hspace{5em}\text{$\triangleright$ commutative monoid $\odot$ with identity $\mymathbb{1}$}\\
a \odot  \mymathbb{1} =  \mymathbb{1} \odot a = a &\\
a \odot b = b \odot a &\\
&\\
a \odot (b\oplus c) = a\odot b \oplus a\odot c  &  \hspace{5em}\text{$\triangleright$ left and right distributive}\\
(a\oplus b) \odot c = a\odot c \oplus b\odot c &\\
&\\
a \odot \mymathbb{0} = \mymathbb{0} \odot a = \mymathbb{0}
\end{align*}
The requirement of being commutative is for the tensor contraction result to be independent of the contraction order, which could be relaxed in certain cases. In the following sections, we show how to compute the independence polynomial,
the maximal independence polynomial, independence number, the number of independent sets, and enumeration of the MIS and maximal independent set configurations of a general graph $G$ by designing tensor element types as commutative semirings while keeping the tensor network generic~\cite{Stepanov2014}.
Table~\ref{tbl:generictypes} summarizes the problems that can be solved by various tensor element types.

@@

Here, we propose to find the independence polynomial by fitting $\alpha(G)+1$ random values of $x_{i}$ and $y_{i} = I(G,x_{i})$. One can then compute the independence polynomial coefficients $a_{i}$ by solving the linear equation: 
\begin{equation}
\left(\begin{matrix}
1 & x_0 & x_0^2 & \ldots & x_0^{\alpha(G)} \\
1 & x_1 & x_1^2 & \ldots & x_1^{\alpha(G)} \\
\vdots & \vdots & \vdots &\ddots & \vdots \\
1 & x_{\alpha(G)} & x_{\alpha(G)}^2 & \ldots & x_{\alpha(G)}^{\alpha(G)}
\end{matrix}\right)
\left(\begin{matrix}
a_0 \\ a_1 \\ \vdots \\ a_{\alpha(G)}
\end{matrix}\right)
= \left(\begin{matrix}
y_0 \\ y_1 \\ \vdots \\ y_{\alpha(G)}
\end{matrix}\right).\label{eq:lineareq}
\end{equation}
With this approach, we do not incur the linear overhead in space. However, because the independence polynomial coefficients can have a huge order-of-magnitude range, if we use floating point numbers in the computation, the round-off errors can be significant for the counting of large independent set sizes.
In addition, the number could easily overflow if we use fixed-width integer types.
The big integer type is also not a good option because big integers with varying width can be very slow and is incompatible with graphics processing unit (GPU) devices. These problems can be solved by introducing a finite-field algebra $\text{GF}(p)$:
\begin{equation}
\begin{split}
    x ~\oplus~ y &= x+y\pmod p,\\
    x ~\odot~ y &= xy\pmod p,\\
    \mymathbb{0} &= 0,\\
    \mymathbb{1} &= 1.
\end{split}\label{eq:finitefield}
\end{equation}
With a finite-field algebra, we have the following observations:
\begin{enumerate}
    \item One can use Gaussian elimination~\cite{Golub2013} to solve the linear equation \Eq{eq:lineareq} since it is a generic algorithm that works for any elements with field algebra. The multiplicative inverse of a finite-field algebra can be computed with the extended Euclidean algorithm.
    \item Given the remainders of a larger unknown integer $x$ over a set of co-prime integers $\{p_1, p_2, \ldots, p_n\}$,
    $x \pmod {p_1 \times p_2 \times \ldots \times p_n}$ can be computed using the Chinese remainder theorem. With this, one can infer big integers from small integers.
\end{enumerate}
With these observations, we develop Algorithm~\ref{alg:finitefield} to compute the independence polynomial exactly without introducing space overheads.
This is an iterative algorithm that tries different prime numbers $p$ until convergence.
In each iteration, we contract the tensor networks to evaluate the polynomial for each $x_{i}$ on the finite field algebra ${\rm GF}(p)$.
Let use denote the results as $Y$, then we solve a linear equation $Y = X A_p$ using the gaussian elimination on ${\rm GF}(p)$ to find the coefficient modulo $p$, $A_p \equiv A \pmod p$.
Finally we apply the Chinese remainder theorem to get $A \pmod {P\times p}$, where $P$ is a product of all prime numbers chosen in previous iterations.
If this number does not change compared with the previous iteration, it indicates the result has converged.
All computations are done with integers of fixed width $W$ except the last step of applying the Chinese remainder theorem.
In \App{app:fft}, we provide another method to solve the linear equation using discrete Fourier transform.

@@

\section{Maximum independent sets and its counting problem}
\subsection{Tropical algebra for finding the independence number and counting MISs}
In the previous section, we focused on computing the independence polynomial for a graph $G$ of given independence number $\alpha(G)$, but we did not show how to compute this number.
The method we use to compute this quantity is based on the following observations. Let $x=\infty$, the independence polynomial becomes
\begin{equation}
I(G, \infty) = a_{\alpha(G)} \infty^{\alpha(G)},
\end{equation}
where the lower-order terms vanish. We can thus replace the polynomial type $a = (a_0, a_1, \ldots, a_k)$ with $a(k) = (a_{k}, k)$, where the second element stores the largest exponent $k$ and the first element stores the corresponding coefficient $a_{k}$.
From this, we can define a new algebra as \red{I've changed this, but I don't have strong feelings about this. We can certainly consider changing back.
I thought this would be more consistent with the earlier notations and show how it's stored in the program as well.}
\blue{I prefer the previous one, what about you @XG?}
\begin{equation}
\begin{split}
    a(k) \oplus a(j) &= \begin{cases}
        \left(a_k + a_j, \max(k,j) \right), & k = j \\
        \left(a_j, \max(k,j) \right), & k < j \\
        \left(a_k, \max(k,j) \right), & k > j
    \end{cases}, \\
    a(k) \odot a(j) &= (a_k a_j, k+j) \\
    \mymathbb{0} &= (0, -\infty)\\
    \mymathbb{1} &= (1, 0). \label{eq:countingtropical}
\end{split}
\end{equation}
%\begin{equation}
%\begin{split}
%    a_x\infty^x \oplus a_y\infty^y &= \begin{cases}
%        (a_x + a_y)\infty^{\max(x,y)}, & x = y\\
%        a_y\infty^{\max(x,y)}, & x < y\\
%        a_x\infty^{\max(x,y)}, & x > y
%    \end{cases}, \\
%    a_x\infty^x \odot a_y\infty^y &= a_x a_y\infty^{x+y}\\
%    \mymathbb{0} &= 0\infty^{-\infty}\\
%    \mymathbb{1} &= 1\infty^{0}. \label{eq:countingtropical}
%\end{split}
%\end{equation}
% In the program, we only need to store the exponent $x$ and the corresponding coefficient $a_x$ initialized to $1$. 
The algebra of the exponents becomes the max-plus tropical algebra: $k \oplus j = \max(k,j)$ and $k \odot j = k + j$~\cite{Maclagan2015, Moore2011}.
This algebra is the same as the one used in Liu et al.~\cite{Liu2021} to calculate and count spin glass ground states.
For independent set calculations here, the vertex tensor and edge tensor becomes:
\begin{equation}
    W^{\rm tropical} = \left(\begin{matrix}
        \mymathbb{1} \\
        (1,1)
    \end{matrix}\right),   
    \qquad \qquad
        B^{\rm tropical} = \left(\begin{matrix}
        \mymathbb{1}  & \mymathbb{1} \\
        \mymathbb{1} & \mymathbb{0}
    \end{matrix}\right).
\end{equation}

@@

This set algebra can serve as the first element in \Eq{eq:countingtropical} to enumerate all MISs, \Eq{eq:polynomial} to enumerate independent sets of different sizes,
or \Eq{eq:max2poly} to enumerate all independent sets of size $\alpha(G)$ and $\alpha(G)-1$.
For example, to enumerate only the MISs, with the tropical algebra, we define $s(k) = (s_{k}, k)$, where the first element follows the algebra in \Eq{eq:set} and the second element follows the max-plus tropical algebra.
The combined operations become: 
\begin{equation}
\begin{split}
    s(k) \oplus s(j) &= \begin{cases}
        \left(s_k \cup s_j, \max(k,j) \right), & k = j \\
        \left(s_j, \max(k,j) \right), & k < j \\
        \left(s_k, \max(k,j) \right), & k > j
    \end{cases}, \\
    s(k) \odot s(j) &= (\{\sigma \lor^\circ \tau \, | \, \sigma \in s_k, \tau \in s_j\}, k+j) \\
    \mymathbb{0} &= (\{ \}, -\infty)\\
    \mymathbb{1} &= (\{0^{\otimes |V|}\}, 0). \label{eq:countingtropicalset}
\end{split}
\end{equation}
%\begin{equation}
%\begin{split}
%    s_x\infty^x \oplus s_y\infty^y &= \begin{cases}
%        (s_x \cup s_y)\infty^{\max(x,y)}, & x = y\\
%        s_y\infty^{\max(x,y)}, & x < y\\
%        s_x\infty^{\max(x,y)}, & x > y
%    \end{cases},\\
%    s_x\infty^x \odot s_y\infty^y &= \{\sigma \lor^\circ \tau | \sigma \in s_x, \tau \in s_y\}\infty^{x+y},\\
%    \mymathbb{0} &= \{\}\infty^{-\infty},\\
%    \mymathbb{1} &= \{0^{\otimes |V|}\}\infty^{0},
%\end{split}
%\end{equation}
Clearly, the vertex tensor and edge tensor become
\begin{equation}
    W^{\rm MISenum}(\{\boldsymbol{e}_{i}\}) = \left(\begin{matrix}
        \mymathbb{1} \\
        \left({\{\boldsymbol{e}_{i}\}, 1} \right)
    \end{matrix}\right),   
    \qquad \qquad
        B^{\rm MISenum} = \left(\begin{matrix}
        \mymathbb{1}  & \mymathbb{1} \\
        \mymathbb{1} & \mymathbb{0}
    \end{matrix}\right).
\end{equation}
The contraction of the corresponding tensor network yields an enumeration of all MIS configurations.

>>>>

\begin{abstract}
We introduce a new approach to compute various properties of independent sets by encoding the problems and the graph into tensor networks with different tensor element algebra.
The type of graph problems that can be computed using this method includes: independence number, the number of independent sets, independence polynomial, maximal independence polynomial, and both counting and enumeration of all maximum independence sets, the large-size independent sets, and all maximal independent sets.
Although the computational complexity inevitably scales exponentially with treewidth of the graph, our method using tensor network contraction with generic programming is highly versatile and is able to compute properties that are otherwise not known to be possible for up to several hundred vertices.
At the same time, our algorithms are very simple to implement and one can directly utilize recent advances in tensor network contraction techniques.
To demonstrate the versatility of this tool, we apply it to a few examples including the calculations of the hard-square entropy constant, Euler characteristics of the independence complex, partition functions, and finite-temperature phase transitions on square-lattice graphs.
\end{abstract}

@@

The independent set problem is closely related to the clique problem and the vertex cover problem~\cite{Moore2011};
more concretely, an MIS of a graph $G$ corresponds to a maximum clique in the complement graph of $G$,
and the complement vertex set of an MIS corresponds to a minimum vertex cover of the graph $G$.
Together, these problems find a wide range of applications in scheduling, logistics, social network analysis, bioinformatics, wireless networks and telecommunication,
map labelling, computer vision, etc.~\cite{Butenko2003, Wu2015}.
Besides the standard problem of finding the independence number and an MIS, there are many other interesting problems pertaining to independent sets, such as the number of independent sets,
independence polynomial, maximal independence polynomial, and enumeration of MIS or maximal independent sets.
Some of these problems are of great interest in physics applications such as the hard-core lattice gas model~\cite{Dyre2016, Fernandes2007}
in statistical mechanics and the Rydberg hamiltonian with neutral atoms~\cite{Pichler2018} \red{cite experiment when it's ready};
they can, for example, be used to understand phase transitions and to identify harder graphs in an ensemble of graphs \red{cite experiment}.

@@

\begin{example}
$C_{ijk} = A_{jkm} B_{mil} V_{jm}$ is a tensor network representing $C_{ijk} = \sum_{ml}A_{jkm} B_{mil} V_{jm}$,
where $i,j,k$ in the output tensor are open edges (note that $l$ is not an open edge and thus summed over).
Its hypergraph representation is shown below, where we use different colors to annotate different hyperedges.

@@

This motivates us to think about what is the most general element type that is allowed in a tensor network contraction program.
We find that as long as the algebra of tensor elements forms a commutative semiring, the tensor network contraction result will be valid and be independent of the contraction order.
A commutative semiring is a semiring with its multiplication operation being commutative, while a semiring is a ring without additive inverse.
To define a commutative semiring with the addition operation $\oplus$ and the multiplication operation $\odot$ on a set $R$, the following relations must hold for any arbitrary three elements $a, b, c \in R$.
\begin{align*}
(a \oplus b) \oplus c = a \oplus (b \oplus c) & \hspace{5em}\text{$\triangleright$ commutative monoid $\oplus$ with identity $\mymathbb{0}$}\\
a \oplus \mymathbb{0} = \mymathbb{0} \oplus a = a &\\
a \oplus b = b \oplus a &\\
&\\
(a \odot b) \odot c = a \odot (b \odot c)  &   \hspace{5em}\text{$\triangleright$ commutative monoid $\odot$ with identity $\mymathbb{1}$}\\
a \odot  \mymathbb{1} =  \mymathbb{1} \odot a = a &\\
a \odot b = b \odot a &\\
&\\
a \odot (b\oplus c) = a\odot b \oplus a\odot c  &  \hspace{5em}\text{$\triangleright$ left and right distributive}\\
(a\oplus b) \odot c = a\odot c \oplus b\odot c &\\
&\\
a \odot \mymathbb{0} = \mymathbb{0} \odot a = \mymathbb{0}
\end{align*}
The requirement of being commutative is for the tensor contraction result to be independent of the contraction order, which could be relaxed in certain cases. In the following sections, we show how to compute the independence polynomial,
the maximal independence polynomial, independence number, the number of independent sets, and enumeration of the MIS and maximal independent sets of a general graph $G$ by designing tensor element types as commutative semirings while keeping the tensor network generic~\cite{Stepanov2014}.
Table~\ref{tbl:generictypes} summarizes the problems that can be solved by various tensor element types.

@@

Here, we propose to find the independence polynomial by fitting $\alpha(G)+1$ random values of $x_{i}$ and $y_{i} = I(G,x_{i})$. One can then compute the independence polynomial coefficients $a_{i}$ by solving the linear equation: 
\begin{equation}
\left(\begin{matrix}
1 & x_0 & x_0^2 & \ldots & x_0^{\alpha(G)} \\
1 & x_1 & x_1^2 & \ldots & x_1^{\alpha(G)} \\
\vdots & \vdots & \vdots &\ddots & \vdots \\
1 & x_{\alpha(G)} & x_{\alpha(G)}^2 & \ldots & x_{\alpha(G)}^{\alpha(G)}
\end{matrix}\right)
\left(\begin{matrix}
a_0 \\ a_1 \\ \vdots \\ a_{\alpha(G)}
\end{matrix}\right)
= \left(\begin{matrix}
y_0 \\ y_1 \\ \vdots \\ y_{\alpha(G)}
\end{matrix}\right).\label{eq:lineareq}
\end{equation}
With this approach, we do not incur the linear overhead in space. However, because the independence polynomial coefficients can have a huge order-of-magnitude range, if we use floating point numbers in the computation, the round-off errors can be significant for the counting of large independent set sizes.
In addition, the number could easily overflow if we use fixed-width integer types.
The big integer type is also not a good option because big integers with varying width can be very slow and is incompatible with graphics processing unit (GPU) devices. These problems can be solved by introducing a finite-field algebra $\text{GF}(p)$:
\begin{equation}
\begin{split}
    x ~\oplus~ y &= x+y\pmod p,\\
    x ~\odot~ y &= xy\pmod p,\\
    \mymathbb{0} &= 0,\\
    \mymathbb{1} &= 1.
\end{split}\label{eq:finitefield}
\end{equation}
With a finite-field algebra, we have the following observations:
\begin{enumerate}
    \item One can use Gaussian elimination~\cite{Golub2013} to solve the linear equation \Eq{eq:lineareq} since it is a generic algorithm that works for any elements with field algebra. The multiplicative inverse of a finite-field algebra can be computed with the extended Euclidean algorithm.
    \item Given the remainders of a larger unknown integer $x$ over a set of co-prime integers $\{p_1, p_2, \ldots, p_n\}$,
    $x \pmod {p_1 \times p_2 \times \ldots \times p_n}$ can be computed using the Chinese remainder theorem. With this, one can infer big integers from small integers.
\end{enumerate}
With these observations, we develop Algorithm~\ref{alg:finitefield} to compute the independence polynomial exactly without introducing space overheads.
This is an iterative algorithm that tries different prime numbers $p$ until convergence.
In each iteration, we contract the tensor networks to evaluate the polynomial for each $\chi_{i}$ on the finite field algebra ${\rm GF}(p)$.
Let use denote the results as $Y$, then we solve a linear equation $Y = X A_p$ using the gaussian elimination on ${\rm GF}(p)$ to find the coefficient modulo $p$, $A_p \equiv A \pmod p$.
Finally we apply the Chinese remainder theorem to get $A \pmod {P\times p}$, where $P$ is a product of all prime numbers chosen in previous iterations.
If this number does not change compared with the previous iteration, it indicates the result has converged.
All computations are done with integers of fixed width $W$ except the last step of applying the Chinese remainder theorem.
In \App{app:fft}, we provide another method to solve the linear equation using discrete Fourier transform.

@@

\section{Maximum independent sets and its counting problem}
\subsection{Tropical algebra for finding the independence number and counting MISs}
In the previous section, we focused on computing the independence polynomial for a graph $G$ of given independence number $\alpha(G)$, but we did not show how to compute this number.
The method we use to compute this quantity is based on the following observations. Let $x=\infty$, the independence polynomial becomes
\begin{equation}
I(G, \infty) = a_{\alpha(G)} \infty^{\alpha(G)},
\end{equation}
where the lower-order terms vanish. We can thus replace the polynomial type $a = (a_0, a_1, \ldots, a_k)$ with $a(k) = (a_{k}, k)$, where the second element stores the largest exponent $k$ and the first element stores the corresponding coefficient $a_{k}$.
From this, we can define a new algebra as
%\begin{equation}
%\begin{split}
%    a(k) \oplus a(j) &= \begin{cases}
%        \left(a_k + a_j, \max(k,j) \right), & k = j \\
%        \left(a_j, \max(k,j) \right), & k < j \\
%        \left(a_k, \max(k,j) \right), & k > j
%    \end{cases}, \\
%    a(k) \odot a(j) &= (a_k a_j, k+j) \\
%    \mymathbb{0} &= (0, -\infty)\\
%    \mymathbb{1} &= (1, 0). \label{eq:countingtropical}
%\end{split}
%\end{equation}
\begin{equation}
\begin{split}
    a_x\infty^x \oplus a_y\infty^y &= \begin{cases}
        (a_x + a_y)\infty^{\max(x,y)}, & x = y\\
        a_y\infty^{\max(x,y)}, & x < y\\
        a_x\infty^{\max(x,y)}, & x > y
    \end{cases}, \\
    a_x\infty^x \odot a_y\infty^y &= a_x a_y\infty^{x+y}\\
    \mymathbb{0} &= 0\infty^{-\infty}\\
    \mymathbb{1} &= 1\infty^{0}. \label{eq:countingtropical}
\end{split}
\end{equation}
In the program, we only need to store the exponent $x$ and the corresponding coefficient $a_x$ initialized to $1$. 
The algebra of the exponents becomes the max-plus tropical algebra: $x \oplus y = \max(x,y)$ and $x \odot y = x + y$~\cite{Maclagan2015, Moore2011}.
This algebra is the same as the one used in Liu et al.~\cite{Liu2021} to calculate and count spin glass ground states.
For independent set calculations here, the vertex tensor and edge tensor becomes:
\begin{equation}
    W^{\rm tropical} = \left(\begin{matrix}
        \mymathbb{1} \\
        (1,1)
    \end{matrix}\right),   
    \qquad \qquad
        B^{\rm tropical} = \left(\begin{matrix}
        \mymathbb{1}  & \mymathbb{1} \\
        \mymathbb{1} & \mymathbb{0}
    \end{matrix}\right).
\end{equation}

@@

This set algebra can serve as the coefficients in \Eq{eq:countingtropical} to enumerate all MISs, \Eq{eq:polynomial} to enumerate independent sets of different sizes,
or \Eq{eq:max2poly} to enumerate all independent sets of size $\alpha(G)$ and $\alpha(G)-1$.
For example, to enumerate only the MISs, with the tropical algebra, we define $s(k) = (s_{k}, k)$, where the first element follows the algebra in \Eq{eq:set} and the second element follows the max-plus tropical algebra.
The combined operations become: 
%\begin{equation}
%\begin{split}
%    s(k) \oplus s(j) &= \begin{cases}
%        \left(s_k \cup s_j, \max(k,j) \right), & k = j \\
%        \left(s_j, \max(k,j) \right), & k < j \\
%        \left(s_k, \max(k,j) \right), & k > j
%    \end{cases}, \\
%    s(k) \odot s(j) &= (\{\sigma \lor^\circ \tau \, | \, \sigma \in s_k, \tau \in s_j\}, k+j) \\
%    \mymathbb{0} &= (\{ \}, -\infty)\\
%    \mymathbb{1} &= (\{0^{\otimes |V|}\}, 0). \label{eq:countingtropicalset}
%\end{split}
%\end{equation}
\begin{equation}
\begin{split}
    s_x\infty^x \oplus s_y\infty^y &= \begin{cases}
        (s_x \cup s_y)\infty^{\max(x,y)}, & x = y\\
        s_y\infty^{\max(x,y)}, & x < y\\
        s_x\infty^{\max(x,y)}, & x > y
    \end{cases},\\
    s_x\infty^x \odot s_y\infty^y &= \{\sigma \lor^\circ \tau | \sigma \in s_x, \tau \in s_y\}\infty^{x+y},\\
    \mymathbb{0} &= \{\}\infty^{-\infty},\\
    \mymathbb{1} &= \{0^{\otimes |V|}\}\infty^{0}. \label{eq:countingtropicalset}
\end{split}
\end{equation}
Clearly, the vertex tensor and edge tensor become
\begin{equation}
    W^{\rm MISenum}(\{\boldsymbol{e}_{i}\}) = \left(\begin{matrix}
        \mymathbb{1} \\
        \left({\{\boldsymbol{e}_{i}\}, 1} \right)
    \end{matrix}\right),   
    \qquad \qquad
        B^{\rm MISenum} = \left(\begin{matrix}
        \mymathbb{1}  & \mymathbb{1} \\
        \mymathbb{1} & \mymathbb{0}
    \end{matrix}\right).
\end{equation}
The contraction of the corresponding tensor network yields an enumeration of all MIS configurations.
